---
title: "Using ECS sample to run matilda"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

The goal of this script is to run Matilda with each of the ECS distributions we sampled prior. 

```{r}
library(matilda)
options(matilda.verbose = FALSE)
library(parallel)
```

# Using ECS samples to run Matilda

We use ECS values sampled from the estimated parameteric distributions from S20 to propagate the varying levels of uncertainty associated with evidence configurations to probabilistic climate projections. This provides an oppotunity to better understand how different ECS evidence configurations affect temperature tragectories from a simple carbon cycle climate model. 

We use the SSP2-4.5 emission scenario to run the model with the perturbed ECS samples.

```{r}
# read in ecs samples as a list
ecs_sample_list <- as.list(readRDS("data/ecs_samples_from_gamma_dist.RDS"))

# read in scenario input file (ini)
ini_245 <- system.file("input/hector_ssp245.ini", package = "hector")

```

The scenario input file is used to initaite a `core` environment for Hector. Hector is the simple carbon-cycle climate model that is the engine behind Matildas prbabilistic projection framework. More details about Hector and its useage, visit the [Hecotor GitHub page](https://jgcri.github.io/hector/).

```{r}
# initiate model core
core <- newcore(ini_245)

```

The result will be a new core object that can will be a required input to run the model. 

# Generate values for other model parameters

Matilda works by running Hector multiple times to build a perturbed parameter ensemble, thus applying parameter uncertainty to model outputs. We need to produce parameter values to accompany the ECS values we sampled in previous steps of the workflow.

Parameter sets are generated in Matilda using `generate_params`. We use this function to produce `n` initial parameter sets (`init_params`).  

```{r}
# set seed for reproducible result
set.seed(123)

# sample size (should match ECS sample)
n = 15000

# generate parameters
init_params <- generate_params(core = core, draws = n)

```

The result will be a new data frame object with 15,000 samples for 6 parameters.

*NOTE*: This data frame includes a column for `ECS`. These are samples drawn from the default prior distribution in Matilda, not the distributions selected for this analysis. 

We replace the default generated `ECS` values with the values we sampled from S20 distributions. This gives us a set of model parameters that are identical except for the `ECS` column, which isolates the impact of propagating `ECS` uncertainty through the model. 

```{r}
# create a list of parameter data frames based on ECS samples in ecs_samnple_list
parameter_list <- lapply(ecs_sample_list, function(data) {
  
  # copy init_params
  params_no_ecs <- init_params
  
  # remove the ECS column from the parameter data frame
  params_no_ecs$ECS <- NULL
  
  # add sampled S20 ecs values
  cbind(data, params_no_ecs)
  
})

# run a check by inspecting a data frame
test <- list(parameter_list$Baseline)
```

