---
title: "Chi-square test on warming probabilities of ECS scenarios"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Goal

The goal of this script is to run a chi-squared test to determine if there is an association between ECS scenario and the likelihood of specific temperature ranges. For example, this can inform whether one ECS configuration leads to a higher likelihood of warming by a given temperature range compared to another ECS configuration. In this analysis, high chi-squared values (significant p-value) suggest that certain ECS configurations influence the likelihood of the temperature range outcome. Conversely, low chi-square statistic (non-significant p-value) suggests there is not enough evidence to conclude that ECS configuration affects the likelihood of a particular warming range. Thus, this analysis can help discern whether ECS configurations have a notable impact on projected temperature outcomes and how the ECS parameter affects future temperature projections.

```{r}
library(reshape2)
library(tidyverse)
```

# Data

This analysis requires the `probability_results`. We load this data from the `data` directory.

```{r}
# load probability data
options(scipen = 999)
prob_data <- readRDS("data/probability_results.RDS") %>%
  rownames_to_column() %>% 
  select(-rowname) %>% 
  mutate(scenario = factor(scenario),
         probability= probability * 100) %>% 
  rename(temperature_range = "bins")

```

Upon loading the data we remove `rownames` and convert `scenario` to a factor.

# Run Chi-square test

H0: There is no significant association between ECS configurations and the likelihood of different end-of-century warming ranges. The probability of temperature ranges is independent of the ECS range.

H1: There is a significant association between ECS configurations and the likelihood of different end-of-century warming ranges. Certain ECS configurations influence the likelihood of certain end-of-century warming occurring. 

```{r}
# Reshape the data into a contingency table format
contingency_table <- dcast(prob_data, 
                           temperature_range ~ scenario, 
                           value.var = "probability")

# # Compute chi-squared results
# result <- fisher.test(contingency_table[-1], simulate.p.value = T)

# split the data by temperature range 
table_split <- split(contingency_table, 
                     contingency_table$temperature_range)

# Run chi-square test on each temperature range
result <- lapply(table_split, function(table) {
  
  # Isolate the probability values in the contingency table
  values <- table[-1]
  
  # Compute chi-squared results
  result <- chisq.test(values)
  
  # Build a results data frame
  stats <- data.frame(X.squared = result$statistic,
                      df = result$parameter,
                      p.value_chi = result$p.value)
  
  return(result)
})

stats_table <- do.call(rbind, result)

```

```{r}
# Visualize the data with a bar plot
library(ggplot2)

# Assuming your data is named 'df'
# Reshape the data from wide to long format
df_long <- pivot_longer(contingency_table, -bins, names_to = "scenario", values_to = "probability")

# Create the bar plot
ggplot(prob_data, aes(x = bins, y = probability, color = scenario, group = scenario)) +
  geom_point()+
  geom_line() +
  labs(title = "Probability Distribution by Scenario",
       x = "Bins",
       y = "Probability") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = c("Baseline" = "blue", 
                                "Baseline_Emergent_constraints" = "green", 
                                "No_Historical" = "red",
                                "No_Paleoclimate" = "orange",
                                "No_Process" = "purple"))
```

